### 2022-01-20 Latex/CoverLetter.tex

% request for removal from telemarketing lists

\documentclass[11pt]{letter}

\usepackage{setspace}
\usepackage{hyperref}
\usepackage{pdfcomment}
\usepackage{gensymb}

\oddsidemargin=.2in
\evensidemargin=.2in
\textwidth=5.9in
\topmargin=-.5in
\textheight=9in

\onehalfspacing
\name{Paul W. Leu, Ph.D.\\
Associate Professor\\
University of Pittsburgh}

\date{\today}

\begin{document}

%\begin{letter}{Prof.~Ali Javey\\ Conexant Systems Distinguished Professor\\ Department of Electrical Engineering \& Computer Sciences\\ University of California, Berkeley}
\begin{letter}{}
%\address{}

%\opening{Dear Editors:}
\opening{Dear Editor:}
 We are enclosing an electronic copy of \textit{Reactive Silver Inks for Antiviral, Repellent Medical Textiles with Ultrasonic Bleach Washing Durability Compared to Silver Nanoparticles}
%by  Sajad Haghanifar, Ping Lu, Md Imrul Kayes, Susheng Tan, Ki-Joong Kim, Tongchuan Gao, Paul Ohodnicki and Paul W. Leu
for publication in %\textit{ACS Applied Bio Materials} 
%\textit{ACS Applied Materials and Interfaces} 
\textit{ACS Applied Nano Materials} as a research article.
%, as recommended by ACS Publications.

% There is a global need for improvements on the anti-viral functionality of reusable personal protective equipment (PPE). Current healthcare protocols in the US require that reusable PPE equipment is washed in bleach solution before being reused. Silver is a popular coating material to enhance the anti-microbial properties of fabric. Many silver fabric treatments demonstrate anti-microbial properties; however, these fabric treatments have not been tested against infectious virus and/or after multiple bleach washing cycles, which is necessary for reusable, functional PPE applications. This work studies the durability of the anti-viral performance for reactive silver ink and silver nanoparticles as viable fabric coatings for reusable, functional PPE applications.

In this paper, we demonstrate a bleach wash durable 
fabric treatment that simultaneously exhibits liquid repellent and anti-viral properties with enveloped virus stocks by using a reactive silver ink and low surface energy polymer. %, even after extended ultrasonic bleach washing. 
The fabric treatment 
exhibits high static contact angles and low contact angle hysteresis with water, even after 300 minutes of ultrasonic bleach washing. Similarly, the fabric treatment shows reductions of infectious virus quantities by about 2 logs compared to controls for enveloped viruses even after bleach washing. The use of reactive silver ink provides for 
better efficacy and durability compared to silver nanoparticles due to the use of reactive molecular silver for more conformal coverage of fabric microfibers and better adhesion.  This study provides insights for improving the wash durability of anti-viral silver fabric treatments and demonstrates a bleach wash durable, repellent anti-viral treatment for reusable, functional PPE applications. 
%The fabricated re-entrant microcell array can find potential applications in various industries including but not limited to electronic devices, medical devices, and optoelectronics.

We submit this manuscript for your consideration of publication in %\textit{ACS Applied Bio Materials} 
%\textit{ACS Applied Materials and Interfaces}.
\textit{ACS Applied Nano Materials}.
Please find a list of experts below as our suggested reviewers.
Thank you for your consideration.
%Thank you.

\closing{Kind Regards,}
%Thank you.
%\closing{Kind Regards,}

\end{letter}

\newpage

\textbf{Submitting Authors:}
%Sajad Haghanifar\\
%Benedum Hall, Room SB10 \\
%Pittsburgh, PA 15261\\
%Email:  \href{mailto:sah175@pitt.edu}{sah175@pitt.edu}\\
%Phone: (412) 499-0768

Anthony J. Galante\\
University of Pittsburgh \\
Department of Industrial Engineering \\
Benedum Hall, Room SB10 \\
Pittsburgh, PA 15261\\
Email:  \href{mailto:ajg109@pitt.edu}{ajg109@pitt.edu}\\

Brady C. Pilsbury \\
University of Pittsburgh \\
Department of Industrial Engineering \\
Benedum Hall, Room SB10 \\
Pittsburgh, PA 15261\\

Kathleen A. Yates\\
University of Pittsburgh Medicine\\
Department of Ophthalmology, Charles T. Campbell Laboratory for Ophthalmic Microbiology\\
Pittsburgh, PA 15213, USA\\

Melbs LeMieux \\
Electroninks Inc \\ 
7901 East Riverside Drive, Bldg 1,Unit 150\\
Austin TX  78744 \\

Daniel J. Bain \\
University of Pittsburgh \\
Department of Geology and Environmental Science\\
Benedum Hall, Room SB10 \\
Pittsburgh, PA 15261\\

Eric G. Romanowski\\
%\altaffiliation{A shared footnote}
University of Pittsburgh Medicine\\
Department of Ophthalmology, Charles T. Campbell Laboratory for Ophthalmic Microbiology\\
Pittsburgh, PA 15213, USA\\
%\phone{+123 (0)123 4445556}
%\fax{+123 (0)123 4445557}

Robert M. Q. Shanks\\
University of Pittsburgh Medicine\\
Department of Ophthalmology, Charles T. Campbell Laboratory for Ophthalmic Microbiology\\
Pittsburgh, PA 15213, USA\\
%Anthony John Galante
 %Mehdi Zarei\\
%Benedum Hall, Room SB10 \\
%Pittsburgh, PA 15261\\
%Email:  \href{mailto:sah175@pitt.edu}{ajg109@pitt.edu}\\
%Phone: (412) 499-0768
%Ping Lu\\
%National Energy Technology Laboratory, U.S. Department of Energy, Pittsburgh, PA 15236, USA\\
%Email:  \href{mailto:Ping.Lu@netl.doe.gov }{Ping.Lu@netl.doe.gov }
%Phone: (412) 499-0768\\

%Anthony John Galante
%Jun Chen\\
%Department of Electrical and Computer Engineering and Peterson Institute of NanoScience and Engineering,\\
%University of Pittsburgh, Pittsburgh, PA 15261, USA \\

%Email:  \href{mailto:sah175@pitt.edu}{ajg109@pitt.edu}\\
%Phone: (412) 499-0768
%Ping Lu\\
%National Energy Technology Laboratory, U.S. Department of Energy, Pittsburgh, PA 15236, USA\\
%Email:  \href{mailto:Ping.Lu@netl.doe.gov }{Ping.Lu@netl.doe.gov }
%Phone: (412) 499-0768\\

%Susheng Tan\\
%Department of Electrical and Computer Engineering and Peterson Institute of NanoScience and Engineering,\\
%University of Pittsburgh, Pittsburgh, PA 15261, USA \\

Paul W. Leu\\
University of Pittsburgh \\
Benedum Hall, Room 1035 \\
Pittsburgh, PA 15261\\
Email:  \href{mailto:pleu@pitt.edu}{pleu@pitt.edu}\\
%Phone: (412) 624-9834

\textbf{Corresponding Author:}\\
Paul W. Leu\\
University of Pittsburgh \\
Benedum Hall Room 1035 \\
Pittsburgh, PA, 15261\\
Email:  \href{mailto:pleu@pitt.edu}{pleu@pitt.edu}\\
Phone: (412) 624-9834

%\pdfcomment[avatar=Paul]{Pick some reviewers from papers we cited.}

%\textbf{Suggested Reviewers:}
\newpage
\textbf{Suggested Reviewers:}
\begin{enumerate}

\item Candace Su-Jung Tsai \\
Associate Professor \\
Department of Environmental Health Sciences, University of California\\
%Phone: +852-2358-8027\\
Email: \href{mailto:candacetsai@ucla.edu}{candacetsai@ucla.edu}\\
%URL: \url{https://www.mcg.msm.cam.ac.uk/people/AL/Ali}\\

%\item Ali Rexa Kamali \\
%Senior Research Associate\\
%Department of Materials Chemistry, University of Cambridge\\
%Phone: +852-2358-8027\\
%Email: \href{mailto:ali@smm.neu.edu.cn}{ali@smm.neu.edu.cn}\\
%URL: \url{https://www.mcg.msm.cam.ac.uk/people/AL/Ali}\\

\item Kostas Kostarelos\\
Professor and Chair\\
Nanomedicine Lab National Graphene Institute and Faculty of Biology Medicine and Health, The University of Manchester\\
Email: \href{mailto:kostas.kostarelos@manchester.ac.uk}{kostas.kostarelos@manchester.ac.uk}\\
%Phone: +86-10-82648147\\
URL: \url{http://www.nanomedicinelab.com/people-and-purpose-2/people/faculty/prof-kostas-kostarelos/}\\

\item Guijun Li\\
Assistant Professor\\
Department of Industrial and Systems Engineering, The Hong Kong Polytechnic University\\
Email: \href{mailto:mitch.li@polyu.edu.hk}{mitch.li@polyu.edu.hk}\\
%Phone: (852) 3469 2958\\
URL: \url{https://seng.hkust.edu.hk/about/people/faculty/mitch-guijun-li}\\

\item Majid Montazer\\
Associate Professor \\
Textile Department, Amirkabir University of Technology\\
Email:  \href{mailto:tex5mm@aut.ac.ir}{tex5mm@aut.ac.ir}\\
%Phone: +86 10 62797530\\
%URL: \url{http://ioe.pim.tsinghua.edu.cn/faculty/baibf/Bai_Eng.htm} \\

\item Eui S. Lee\\
Assistant Professor \\
Department of Textile Engineering, Inha University\\
Email: \href{mailto:eslee@inha.ac.kr}{eslee@inha.ac.kr}\\
%Phone: +852-3442 4897 \\
%URL: \url{http://www.ap.cityu.edu.hk/personal-website/johnny/site_flash/index.html}\\

\item Mh Busra Fauzi\\
Assistant Professor \\
Centre for Tissue Engineering and Regenerative Medicine, University Kebangsaan Malaysia  \\
Email: \href{mailto:fauzibusra@ukm.edu.my}{fauzibusra@ukm.edu.my}\\
%%Phone: (886)3-5715131 ext 33965 \\
%URL: \url{https://rtc.sandia.gov/the-rtc-team/}

%\item Yu-Lun Chueh\\
%Associate Professor\\
%Material Science and Engineering, National Tsing Hua University \\
%Email: \href{mailto:ylchueh@mx.nthu.edu.tw}{ylchueh@mx.nthu.edu.tw}\\
%Phone: (886)3-5715131 ext 33965 \\
%URL: \url{http://nanoscienceandnanodevicelab.weebly.com/advisor.html}
\end{enumerate}

%\begin{enumerate}
%
%
%%URL: \url{http://www.ap.cityu.edu.hk/personal-website/johnny/site_flash/index.html}\\
%
%%

%
%
%\item Bharat Bhushan\\
%
%Professor in the Department of Mechanical and Aerospace Engineering, The Ohio State University\\
%Email: \href{mailto:bhushan.2@osu.edu}{bhushan.2@osu.edu}\\
%%Phone: +31-20-7547100 \\
%URL: \url{https://mae.osu.edu/people/bhushan.2}
%
%
%%\item Liangbing Hu\\
%
%%Assistant Professor\\
%%Professor of Photonic Materials for Photovoltaics, University of Amsterdam\\
%%Email: \href{mailto: binghu@umd.edu}{ binghu@umd.edu}\\
%%Phone: +1-301-405-9303 \\
%%URL: \url{http://bingnano.com/}
%
%\end{enumerate}

\end{document}